---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE}
options(tibble.print_min = 5)
```

# Data {#data}

## Getting the data {#retrieve}

The data is retrieved from the Federal Reserve website, under the discussion
series: _The U.S. Treasury Yield Curve: 1961 to the Present_. The link for that
site is [here](https://www.federalreserve.gov/pubs/feds/2006/200628/200628abs.html).
The specific data set downloaded was the XLS file included on that site. `ratekit` provides the 
`download_rates_xls()` helper function for this.

The data was immediately opened in Excel, and resaved as an `xlsx` file.
The format of the raw data is not a true `xls` file, rather, it is some flavor of `xml`
file. This does not play nicely with R's packages for importing Excel data, so
a resave was necessary and is done manually.

## Cleaning {#cleaning}

Data is brought in using the `readxl` package and the `ratekit` helper, `read_rates()`.
This function reads the rectangle of rates data only, and sets any `-999.99` values
to `NA`. These are often found through the dataset, and it is assumed that they represent missing values.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(skimr)
library(ratekit)
library(dplyr)
library(tibbletime)
library(readr)

raw <- read_rates("data/raw/feds200628.xlsx")

# Only select the parameters
raw_params <- select(raw, date, contains("BETA"), contains("TAU"))
```

The column names in the data correspond to different types and lengths of rates used in the 
paper. The key for understanding the column names is below:

```{r meta, echo=FALSE}
meta <- readRDS("data/cleaned/meta/meta.rds")

meta %>%
  rename(Series = series, `Compounding Convention` = compounding_convention, `Key` = key) %>%
  knitr::kable(booktabs = TRUE, caption = "Rates data: Column key") %>% 
  kableExtra::kable_styling(position = "center", latex_options = "HOLD_position")
```

Most of these columns are not important for this analysis. Only the parameter
columns and the date column are kept. To further examine the missing values,
the `skimr` package was used, producing the following report.

```{r par_yield_missing, echo=FALSE, results='asis'}
params <- select(raw, date, contains("BETA"), contains("TAU"))

skim_with(numeric = list(hist = NULL), ts = list(line_graph = NULL))
skim(params) %>% 
  kable(booktabs = TRUE, caption = "`skimr` summary statistics for the parameters")
```

The `TAU2` column has a number of missing values. All of them occur before 1980,
and were removed from the data set. After that removal, no missing values remain.

## Monthly and Ascending {#monthly}

```{r, echo=FALSE}
parameters <- read_rds("data/cleaned/parameters/parameters.rds")

parameters_monthly <- parameters %>%
  as_tbl_time(date) %>%
  arrange(date) %>% 
  as_period("monthly", side = "end")
```

Monthly data is required for the report, but daily data is provided from the Fed.
The data is converted to monthly using the `tibbletime` package. This leaves `r nrow(parameters_monthly)` rows of data for the project, spanning `r min(parameters_monthly$date)` to `r max(parameters_monthly$date)`.