---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE}
options(tibble.print_min = 5)
```

# Data {#data}

## Data Retrieval {#retrieve}

The data is retrieved from the Federal Reserve website, under the discussion
series: _The U.S. Treasury Yield Curve: 1961 to the Present_. The link for that
site is: [https://www.federalreserve.gov/pubs/feds/2006/200628/200628abs.html](https://www.federalreserve.gov/pubs/feds/2006/200628/200628abs.html).
The specific data set downloaded was the XLS file included on the site. `ratekit` provides the 
`download_rates_xls()` helper function for this.

The data was immediately opened in Excel, and resaved as an `xlsx` file.
The format of the raw data is not a true `xls` file, rather, it is some flavor of an `xml`
file. This does not play nicely with R's packages for importing Excel data, so
a resave was necessary and is done manually.

## Cleaning {#cleaning}

Data is brought into R using the `readxl` package and the `ratekit` helper, `read_rates()`.
This function sets any `-999.99` values to `NA`. 
These are often found through the dataset, especially in the parameters columns,
and it is assumed that they represent missing values.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(skimr)
library(ratekit)
library(dplyr)
library(tibbletime)
library(readr)

raw <- read_rates("data/raw/feds200628.xlsx")

# Only select the parameters
raw_params <- select(raw, date, contains("BETA"), contains("TAU"))
```

The column names in the data correspond to different types and lengths of rates used in the 
paper, along with the names of the parameters in the model. The key for understanding the column names is
shown in Table \@ref(tab:meta).

```{r meta, echo=FALSE, warning=FALSE}
meta <- readRDS("data/cleaned/meta/meta.rds")

meta %>%
  rename(Series = series, `Compounding Convention` = compounding_convention, `Key` = key) %>%
  knitr::kable(booktabs = TRUE, caption = "Rates data: Column key") %>% 
  kableExtra::kable_styling(position = "center", latex_options = "HOLD_position")
```

Most of these columns are not important for this analysis. Only the parameter
columns and the date column are kept. To further examine the missing values,
the `skimr` package was used, producing the report shown below. The `TAU2` column 
has a number of missing values (resulting from either being missing or from
being `-999.99` values assumed to be missing). All of them occur before 1980,
and were removed from the data set. After that removal, no missing values remain,
and the values for the other parameters seemed to stabilize as well.

```{r par_yield_missing, echo=FALSE, results='asis'}
params <- select(raw, date, contains("BETA"), contains("TAU"))

skim_with(numeric = list(hist = NULL), ts = list(line_graph = NULL))
skim(params) %>% 
  kable(booktabs = TRUE, caption = "`skimr` summary statistics for the parameters")
```


## Monthly and Ascending {#monthly}

```{r, echo=FALSE}
parameters <- read_rds("data/cleaned/parameters/parameters.rds")

parameters_monthly <- parameters %>%
  as_tbl_time(date) %>%
  arrange(date) %>% 
  as_period("monthly", side = "end")
```

Monthly data is required for the report, but daily data is provided from the Federal
Reserve data set. The data is converted to monthly (end-of-month) using the `tibbletime` package. This leaves `r nrow(parameters_monthly)` rows of data for the project, spanning `r min(parameters_monthly$date)` to `r max(parameters_monthly$date)`.