---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE}
options(tibble.print_min = 5)
```

# Data {#data}


## Getting the data {#retrieve}

Script) [01-download.R](./R/01-download.R)

The data is retrieved from the Federal Reserve website, under the discussion
series: _The U.S. Treasury Yield Curve: 1961 to the Present_. The link for that
site is [here](https://www.federalreserve.gov/pubs/feds/2006/200628/200628abs.html).
The specific data set that was downloaded was the XLS file included on that site.

The data was immediately opened in Excel, and was resaved as an `xlsx` file.
The format of the data is not a true `xls` file, instead it is some kind of `xml`
file. This does not play nicely with R's packages for importing Excel data, so
a resave was necessary and is done manually.

`ratekit` provides the `download_rates_xls()` helper function for this.

## Cleaning {#cleaning}

Script) [02-cleaning.R](./R/02-cleaning.R)

Data is brought in using the `readxl` package and the `ratekit` helper, `read_rates()`.
This function reads the rectangle of rates data only, and sets any `-999.99` values
to `NA`. These are often found through the dataset, and I assume they are meant
to represent missing values. To visualize the `NA` values in the dataset, I use
the `visdat` package.

First, let's look at what is immediately brought in by `read_rates()`. We will
need a few packages throughout the chapter, so let's load those now as well.

```{r, message=FALSE, warning=FALSE}
library(visdat)
library(ratekit)
library(dplyr)
library(tibbletime)

raw <- read_rates("data/raw/feds200628.xlsx")

raw
```

Not a bad start, but I'm worried about missing values. Also, what are those
column names?

The column names in the data correspond to different types and lengths of rates used in the 
paper. The key for understanding the column names is below:

```{r meta, echo=FALSE}
meta <- readRDS("data/cleaned/meta/meta.rds")
knitr::kable(meta)
```

Using `vis_dat()`, we can take a look at our dataset all at once to determine
which data points to exclude.

```{r missing_vals, fig.width=10, fig.height=6, cache=TRUE}
vis_dat(raw, warn_large_data = FALSE)
```

A clear pattern is seen in the missing values, with the number of missing values
increasing as you go further back in time and look at longer rates (10 year VS 30 year). 
This might be a bit difficult to see if you look at everything, but becomes clearer
if you zoom in on just one set of series.

```{r par_yield_missing, echo=FALSE, fig.width=10, fig.height=6, cache=TRUE}
par_yld <- select(raw, date, contains("SVENPY"))
vis_dat(par_yld)
```

The parameters are affected by this as well, but not as much, with only `TAU2` 
being affected.

```{r parameters_missing, echo=FALSE, fig.width=10, fig.height=6, cache=TRUE}
params <- select(raw, date, contains("BETA"), contains("TAU"))
vis_dat(params)
```

Since the parameters are all we care about for this project, I decided to throw
out any row with an `NA` value for `TAU2`. This threw out every data
point before 1980. We can ensure that we don't have any missing values now 
with `vis_miss()`.

```{r parameters_clean, fig.width=10, fig.height=6, cache=TRUE}
# This is the cleaned parameter set, cleaned using 02-cleaning.R
parameters <- readRDS("data/cleaned/parameters/parameters.rds")
vis_miss(parameters)
```

## Monthly and Ascending {#monthly}

Script) [03-to-monthly-and-ascending.R](./R/03-to-monthly-and-ascending.R)

At this point, our dataset looks like this:

```{r}
parameters <- readRDS("data/cleaned/parameters/parameters.rds")
parameters
```

We want monthly data, and we will need to put it in ascending order. We can
convert to monthly with `as_period()` from `tibbletime`, and arrange it by
ascending date with `arrange()` from `dplyr`.

```{r}
parameters_monthly <- parameters %>%
  as_tbl_time(date) %>%
  arrange(date) %>% 
  as_period("monthly", side = "end")

parameters_monthly
```

This leaves us with `r nrow(parameters_monthly)` rows of data for our project, 
spanning `r min(parameters_monthly$date)` to `r max(parameters_monthly$date)`.