---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE}
options(tibble.print_min = 5)
```

# Question 2 {#q2}

**Question:**

_Can the three yield curve factors explain the time-series variation in spot rates? Regress $y_t(1)$ on a constant and $X_t$ and comment on the regression statistics. Perform the same analysis for $y_t(5)$ and $y_t(10)$._

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(ratekit)
library(ggplot2)
library(broom)
library(purrr)
library(furrr)
library(rsample)
library(tibbletime)
library(forcats)
library(patchwork)
```

In this question, data for the $y_t(1)$, $y_t(5)$, and $y_t(10)$ spot rates
is required along with the yield curve factors, $X_t$. These have been calculated in Section \@ref(spot-rates) and Section \@ref(yield-curve-factors), and the results from there are used here.

```{r, echo=FALSE}
rates    <- read_rds("data/computed/rates.rds")
yield_curve_factors <- read_rds("data/computed/yield_curve_factors.rds")
n <- c("1", "5", "10")

spot_rates_q2 <- rates
```

## Regression {#regression}

The following regression was run on every spot rate series:

$$ y_{t}(n) = \alpha(n) + \beta_1(n) \text{Level}_t + \beta_2(n) \text{Slope}_t + \beta_3(n) \text{Curvature}_t + \epsilon_t(n) $$

Using the concept of multiple models from the book, [`R 4 Data Science`](http://r4ds.had.co.nz/many-models.html), implementing these regressions in R is incredibly straightforward. The general concept involves two steps:

1) Split the data set of all 14 maturities into 14 groups, stored in a nested data frame.
2) For each group, run the linear model above and store the result.

```{r, echo=FALSE}
nested_spot <- spot_rates_q2 %>%
  # Add on the curve factors
  left_join(yield_curve_factors, by = "date") %>%
  
  # Group by maturity and nest
  group_by(maturity) %>%
  select(-maturity_nm) %>%
  nest()
```

```{r, echo=FALSE}
nested_models <- nested_spot %>%
  mutate(model = map(data, ~ lm(spot_rate ~ level + slope + curvature, data = .x)))
```

The final result is a compact single data frame that contains the 14 resulting
models along with the original data, indexed by the maturity.

```{r, echo=FALSE}
nested_models
```


## One Year Spot Rate 

As seen in Table \@ref(tab:one-year-reg) and Table \@ref(tab:one-year-reg-r2), all estimates for the 1 year spot rate model are highly significant, 
and the Adjusted $R^2$ is nearing 100%, suggesting that the model can explain
essentially all of the variation in the spot rate. Considering that the 
constructed yield curve factors include $y_t(2)$ as the level effect, this
should not be surprising as all of the series are highly correlated. 
The very high statistic on the level coefficient supports the claim of its importance.

```{r one-year-reg, echo=FALSE, warning=FALSE}
model_1_yr <- nested_models %>%
  filter(maturity == 1) %>%
  pluck(list("model", 1))

model_1_yr %>%
  tidy() %>%
  mutate(term = recode(term, "(Intercept)" = "Intercept", "level" = "Level", 
                       "slope" = "Slope", "curvature" = "Curvature")) %>%
  rename(Term = term, Estimate = estimate, `Standard Error` = std.error, 
         `Statistic` = statistic, `P-Value` = p.value) %>%
  knitr::kable(digits = 15, booktabs = TRUE, caption = "Regression results: One year spot") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position", position = "center")
```

```{r one-year-reg-r2, echo=FALSE, warning=FALSE}
model_1_yr %>%
  glance() %>%
  select(`R Squared` = r.squared, `R Squared Adj` = adj.r.squared, `Residual Std Error` = sigma) %>%
  knitr::kable(booktabs = TRUE, caption = "Regression $R^2$: One year spot") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position", position = "center")
```

A chart of the realized VS predicted time series for each model confirms how well the variation
is explained, with essentailly perfect matching of the realized series. 
It is important to remember that this model is not predicting
out-of-sample future rates, and is simply used to gather intuition about past rates. Nevertheless,
it is interesting to see how well the yield curve factors explain the variation.

```{r spot-rate-reg-chart, fig.width=10, echo=FALSE, fig.pos = 'H', fig.cap="In-sample realized VS predicted spot rate series", warning=FALSE, fig.align='center', cache=TRUE}
nested_with_predictions <- nested_models %>%
  mutate(data = map2(data, model, ~mutate(.x, predictions = predict(.y))))
  
nested_with_predictions %>%
  filter(maturity %in% c(1, 5, 10)) %>%
  mutate(maturity = recode(maturity, "1" = "One Year Spot", "5" = "Five Year Spot", "10" = "Ten Year Spot")) %>%
  mutate(maturity = forcats::as_factor(maturity)) %>%
  unnest(data) %>%
  rename(Realized = spot_rate, Predictions = predictions) %>%
  select(-(level:curvature)) %>%
  gather("Realized/Predicted", "Spot Rate", Realized, Predictions) %>%
  ggplot(aes(x = date, y = `Spot Rate`, color = `Realized/Predicted`)) +
  geom_line() +
  facet_wrap(~maturity, nrow = 3) +
  theme_minimal() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "")
```

## Five Year Spot Rate 

The model for the 5-year rate is similar to the 1-year rate in terms of explanatory power.
The level coefficient is slightly higher than for the 1-year, and is >1, which
makes sense considering that the 5-year is generally higher than the 2-year
spot rate used to represent the level effect. The slope is more than triple
that of the 1-year, suggesting that an increase in the slope has a larger impact on the value of the
5-year curve than it does on the 1-year.

```{r five-year-reg, echo=FALSE, warning=FALSE}
model_5_yr <- nested_models %>%
  filter(maturity == 5) %>%
  pluck(list("model", 1))

model_5_yr %>%
  tidy() %>%
  mutate(term = recode(term, "(Intercept)" = "Intercept", "level" = "Level", 
                       "slope" = "Slope", "curvature" = "Curvature")) %>%
  rename(Term = term, Estimate = estimate, `Standard Error` = std.error, 
         `Statistic` = statistic, `P-Value` = p.value) %>%
  knitr::kable(digits = 15, booktabs = TRUE, caption = "Regression results: Five year spot") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position", position = "center")
```

```{r five-year-reg-r2, echo=FALSE, warning=FALSE}
model_5_yr %>%
  glance() %>%
  select(`R Squared` = r.squared, `R Squared Adj` = adj.r.squared, `Residual Std Error` = sigma) %>%
  knitr::kable(booktabs = TRUE, caption = "Regression $R^2$: Five year spot") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position", position = "center")
```

## Ten Year Spot Rate 

Finally, the 10-year model performs similarly to the 1 and 5-year models. One
interesting thing to note about the 10-year model is that the sign on the curvature
coefficient is opposite that of the 1 and 5-year models.

```{r ten-year-reg, echo=FALSE, warning=FALSE}
model_10_yr <- nested_models %>%
  filter(maturity == 10) %>%
  pluck(list("model", 1))

model_10_yr %>%
  tidy() %>%
  mutate(term = recode(term, "(Intercept)" = "Intercept", "level" = "Level", 
                       "slope" = "Slope", "curvature" = "Curvature")) %>%
  rename(Term = term, Estimate = estimate, `Standard Error` = std.error, 
         `Statistic` = statistic, `P-Value` = p.value) %>%
  knitr::kable(digits = 15, booktabs = TRUE, caption = "Regression results: Ten year spot") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position", position = "center")
```

```{r ten-year-reg-r2, echo=FALSE, warning=FALSE}
model_10_yr %>%
  glance() %>%
  select(`R Squared` = r.squared, `R Squared Adj` = adj.r.squared, `Residual Std Error` = sigma) %>%
  knitr::kable(booktabs = TRUE, caption = "Regression $R^2$: Ten year spot") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position", position = "center")
```

## Decomposing the Spot Curve

Although not specifically asked for, it might be interesting to 
decompose and plot the spot curve at a few particular points in time. The procedure for this involves
the following manipulations:

1) For month `m`, extract the spot rate at every available maturity for that month.
2) For month `m`, filter the yield curve factors down to that month, and multiply
each maturity's regression coefficients by the corresponding yield curve factor. This
gives the contribution of each factor for that month and each maturity.
3) Join the two data sets and chart them to view the decomposed spot rate for any month.

```{r, echo=FALSE}
extract_rates <- function(.rates, .date) {
  .rates %>%
    as_tbl_time(date) %>%
    group_by(maturity) %>%
    filter_time(~.date)
}
```

```{r, echo=FALSE}
tidy_models <- function(.nested_models, .date) {
  coefs <- .nested_models %>%
    mutate(coef = map(model, tidy)) %>%
    unnest(coef) %>%
    select(maturity, term, estimate) %>%
    spread(term, estimate) %>%
    select(-`(Intercept)`)
  
  yield_fct <- yield_curve_factors %>%
    as_tbl_time(date) %>%
    filter_time(~.date)
  
  # Multiply the date's yield curve factors by the coefficients to get the
  # decomp of the term
  decomp <- coefs %>%
    mutate(level = level * yield_fct$level,
           slope = slope * yield_fct$slope,
           curvature = curvature * yield_fct$curvature)
  
  decomp
}
```

```{r, echo=FALSE}
tidy_decomposed_spot_rate <- function(.rates, .nested_models, .date) {
  .rates %>% 
    extract_rates(.date) %>% 
    left_join(tidy_models(.nested_models, .date), "maturity") %>%
    gather("line", "value", -(maturity:date)) %>%
    select(maturity, line, value)
}
```

```{r, echo=FALSE}
# A charting function with custom themes
chart_decomposed_spot_rate <- function(.decomposed_spot) {
  .decomposed_spot %>%
    mutate(linetype = case_when(
      line == "Spot Rate" ~ "a", # this works by alphabetical. first is solid, then dashed
      TRUE ~ "b"
    )) %>%
    ggplot(aes(x = maturity, y = value, color = line, linetype = linetype)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE) +
    theme_minimal() +
    scale_color_brewer(palette = "Dark2") +
    labs(x = "Maturity", y = "Component Value", color = "Component") +
    guides(linetype = FALSE)
}
```

This procedure was streamlined into a single function, parameterized by the month
to allow for easy plotting and comparison of multiple months. For example, the 
decomposed spot curve for January 2012 and January 1981 are shown side-by-side
in Figure \@ref(fig:decomposed-spot). January of 1981 was definitely an interesting time period! The spot rate is
essentially inverted, with lower maturity bonds having higher spot rates than
longer maturity bonds.

```{r decomposed-spot, fig.width=10, echo=FALSE, fig.pos = 'H', fig.cap="Decomposed spot rate for January of 2012 and 1981", warning=FALSE, fig.align='center', cache=TRUE}
p1 <- rates %>%
  tidy_decomposed_spot_rate(nested_models, "2012-01") %>%
  mutate(line = recode(line, "spot_rate" = "Spot Rate", "level" = "Level",
                       "slope" = "Slope", "curvature" = "Curvature")) %>%
  chart_decomposed_spot_rate() +
  labs(title = "January 2012") +
  theme(legend.position = "bottom")

p2 <- rates %>%
  tidy_decomposed_spot_rate(nested_models, "1981-01") %>%
  mutate(line = recode(line, "spot_rate" = "Spot Rate", "level" = "Level",
                       "slope" = "Slope", "curvature" = "Curvature")) %>%
  chart_decomposed_spot_rate() +
  labs(title = "January 1981", y = "") +
  theme(legend.position = "bottom")

p1 + p2 + plot_layout(ncol = 2)
```

## Coefficient Stability {#coef-stability}

Another question worth asking is how stable the coefficients are throughout time. We can test this by running the same regression as before, but with a _rolling window_. This works
by calculating the regression `spot_rate ~ level + slope + curvature` for the
first 100 months, then shifting forward 1 month and dropping the last day, calculating the
regression again, and repeating this for the length of the series. The `rsample`
package provides a number of helpers for doing analysis exactly like this. The procedure for this is:

1) Split the data into 14 nested groups by maturity, as done in Section \@ref(regression).
2) Further split each of the 14 groups into 359 rolling subsets of 100 days each using `rolling_origin()` from `rsample`.
3) For each maturity, and for each rolling split, run the regression from \@ref(regression). This results in
5026 regressions, which, on this 4-core computer, can be run in parallel in ~8 seconds.

```{r, echo=FALSE}
all_splits <- spot_rates_q2 %>%
  # Add on the curve factors
  left_join(yield_curve_factors, by = "date") %>%
  group_by(maturity) %>%
  select(-maturity_nm) %>%
  nest() %>%
  # For each maturity, split the data into subsets of 100 data points
  mutate(data_splits = map(data, ~rolling_origin(.x, initial = 100, assess = 1, cumulative = FALSE)))
```

```{r, eval=FALSE, echo=FALSE}
# If we were to dig into just 1 of the `data_splits` we would find another tibble containing the slices. 
all_splits$data_splits[[1]]
```


```{r, echo=FALSE}
# This notation might all look complicated, but it turns out to be incredibly useful
# and scalable. To prove that, let's run the model on every split, for every maturity.

plan(multiprocess)

all_splits <- all_splits %>%
  mutate(
    model_coef = future_map(data_splits, ~{ # For each maturity...
      
      maturity_splits <- .x
      
      map_dfr(maturity_splits$splits, ~ { # For each split...
        
        split.x <- .x
        date <- assessment(split.x)$date

        # Run the model
        mod <- lm(spot_rate ~ level + slope + curvature, data = analysis(split.x))

        # Tidy up
        tidy_mod <- mod %>% tidy()
        mutate(tidy_mod, date = date) %>%
          select(date, term, estimate)
      })
    })
  )

```

When the procedure has been run, a natural next step is to pick some of the 14 maturities
and look at the stability of each coefficient over time. For example, the 1, 5, and 10 year coefficients over time
are shown side-by-side in Figure \@ref(fig:coef-stability). Most of the coefficients are fairly stable
over time, with the exception of curvature. The curvature of the 1 year has begun to rise up from -0.5 to around -0.25 in recent years. This change over time is not refected in the static -0.428 curvature estimate we get from running the model
over the full time period, and might offer other interesting insights or opportunities for trading strategies.
The 10 year curvature coefficient follows a similar pattern,
but the 5 year curvature follows the opposite trend, decreasing in recent years.

```{r, echo=FALSE}
plot_coef_stability <- function(.x, .maturity) {
  coefs <- .x %>%
    filter(maturity == .maturity) %>%
    pull(model_coef) %>%
    pluck(1)
  
  coefs %>%
    mutate(term = as_factor(term)) %>%
    mutate(term = recode(term, `(Intercept)` = "Intercept", curvature = "Curvature", 
                         level = "Level", slope = "Slope")) %>%
    ggplot(aes(x = date, y = estimate)) +
    geom_line() +
    facet_wrap(~term, nrow = 4) +
    labs(x = "", y = "Estimate") +
    theme_minimal() +
    coord_cartesian(ylim = c(-.6, 1.1))
}
```

```{r coef-stability, fig.width=10, echo=FALSE, fig.pos = 'H', fig.cap="Coefficient stability for 1, 5, and 10 year spot rates. 100 month rolling window.", warning=FALSE, fig.align='center', fig.height=8, cache=TRUE}

coef_p1 <- plot_coef_stability(all_splits, 1) + labs(title = "One Year Coefficients")
coef_p2 <- plot_coef_stability(all_splits, 5) + labs(title = "Five Year Coefficients", y = "")
coef_p3 <- plot_coef_stability(all_splits, 10) + labs(title = "Ten Year Coefficients", y = "")

coef_p1 + coef_p2 + coef_p3 + plot_layout(ncol = 3)
```
