---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE}
options(tibble.print_min = 5)
```

# Question 3 {#q3}

**Question:**

_Perform the “out-of-sample” hedging exercises where $n_h = 3$. Split the sample into two halves, such that $T_0 = 2T$. Begin with the $T$-th month, calculate $w_T(1)$, $w_T(5)$, and $w_T(10)$ for each hedging strategy, and save $\epsilon_{T+\Delta}(n_h)$. Move forward and repeat the process each month, and calculate the root mean squared hedging error (RMSHE) for each strategy:_

$$ RMSHE = [\frac{1}{T} \sum{\epsilon_{t+\Delta}(n_h)}]^{0.5} $$ 

_Report your results and evaluate the performance of the hedging strategies._

**Implementation details:**

This chapter is broken into:

- 3 sections devoted to creating the modeling functions
- 1 section for creating a hedging error calculation function
- 1 section for creating an "interface" to the models and error calculations
- 1 section for running the models through the interface
- 1 section for reviewing model performance

```{r, message=FALSE, warning=FALSE, echo = FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(ratekit)
library(ggplot2)
library(broom)
library(purrr)
library(rsample)
library(tibble)
```

The data for the $y_t(1)$, $y_t(3)$, $y_t(5)$, $y_t(7)$, and $y_t(10)$ spot rates is required to calculate the durations for the zero coupon bonds used in the hedging exercises. Also required are the yield curve factors and excess returns. These have all been calculated in Section \@ref(rates), and the results from there are used.

```{r, echo=FALSE}
rates         <- read_rds("data/computed/rates.rds")
factors       <- read_rds("data/computed/yield_curve_factors.rds")
excess_return <- read_rds("data/computed/excess_returns.rds")

# Only these maturities are used
mats <- c("1", "3", "5", "7", "10")

spot_rates_q3 <- rates %>%
  filter(maturity_nm %in% mats)

T_full <- spot_rates_q3 %>% filter(maturity == 1) %>% nrow()
T_half <- round(T_full / 2)

# Half of the number of data points
# Used as the train / test split
# T_half
```
 
The models developed below will be structured in a way so that they can all be
fed with data from the same data set. To accomplish the rolling out-of-sample
technique, the `rsample` package was used. Specifically, the `rolling_origin()` function
was utilized which allows easy splitting of data into rolling sets of `r T_half` months, 
which is the `T` value corresponding to half of the data.

```{r, echo=FALSE}
# First, a full data set was constructed that can be used by all models.

full_data <- spot_rates_q3 %>% 
  left_join(excess_return, c("maturity_nm", "date")) %>% 
  left_join(factors, "date") %>%
  arrange(date)
```


```{r, echo=FALSE}
# By arranging by `date` as done above, it becomes clear that to shift forward 1 month,
# you actually have to shift forward 5 rows, because there are 5 maturities involved in
# this data set. `rolling_origin()` makes this straightforward.

n <- length(mats)

model_data <- full_data %>%
  rolling_origin(initial = T_half * n, assess = 1 * n, cumulative = FALSE, skip = n-1)
```

```{r, eval=FALSE, echo=FALSE}
# We now have `r nrow(model_data)` rolling sets to use in each of our models. Each 
# set is self contained, containing both the data used for training that set (which in this case means determining the weights used for the next month), and calculating the hedging error for the next month.

# Training data
analysis(model_data$splits[[1]]) %>% print(n = 10)
```

```{r, eval=FALSE, echo=FALSE}
# Testing data
assessment(model_data$splits[[1]])
```
 
## Modified / Macaulay Duration

The first two strategies involve creating duration matched barbells using 1-year 
and 10-year zero coupon bonds to match the duration of a 3-year or 7-year bullet.
Since the strategies are essentially the same, one modeling function is used for both, `model_duration()`,
with an argument to select whether to use modified or macaulay duration. The duration
is calculated using the `duration()` function in the `ratekit` package 
and the barbell weights are calculated with the `barbell_weights()` function.

The function is created in such a way that it accepts a single split from the 
total data frame, and calculates the weights for just that one split.
The weight for a 5-year zero is included to be consistent with the other models,
but is set to `0`.

```{r, echo=FALSE}
model_duration <- function(split, bullet = "3", type = "modified") {
  # For modified duration, we only need the most recent month.
  train <- analysis(split) %>% tail(5L)
  
  w_1 <- train %>%
    mutate(mod_duration = duration(n = maturity, c = 0, M = 1, y = spot_rate, type = type)) %>%
    select(date, maturity_nm, mod_duration) %>% 
    spread(maturity_nm, mod_duration) %>%
    select(date, `1`, `10`, bullet) %>%
    gather(bullet_maturity, bullet_duration, bullet) %>%
    mutate(w_1 = barbell_weights(bullet_duration, `1`, `10`)) %>%
    pull(w_1)
    
  tibble(
    maturity_nm = c("1", "5", "10"),
    weight = c(w_1, 0, 1 - w_1)
  )
}
```

## Simple Regression Based Hedging

The regression based hedging method involves the regression:

$$ ER_t(n_h) = w_t(1) ER_t(1) + w_t(5) ER_t(5) + w_t(10) ER_t(10) + u_t(n_h) $$

The R function `lm()` is used to run the regression, and the weights are 
extracted and returned in the same format as the duration models. This is all 
wrapped into a modeling function, `model_regression()`.

```{r, echo=FALSE}
model_regression <- function(split, bullet = "3") {
  train <- analysis(split)
  
  train_data <- train %>%
    select(date, maturity_nm, ER_n) %>% 
    spread(maturity_nm, ER_n)
  
  form <- as.formula(paste0("`", bullet, "` ~ 0 + `1` + `5` + `10`"))
  
  model <- lm(form, data = train_data)
  
  weight_tbl <- model %>%
    coef() %>%
    enframe() %>%
    rename(maturity_nm = name, weight = value) %>%
    mutate(maturity_nm = gsub("`", "", maturity_nm))
    
  weight_tbl
}
```

## Multiplicative Regression Hedging

The multiplicative regression follows the model:

$$ ER_t(n_h) = \theta_t(1; X_{t-\Delta}) ER_t(1) + \theta_t(5; X_{t-\Delta}) ER_t(5) + \theta_t(10; X_{t-\Delta}) ER_t(10) + u_t(n_h) $$
where $\theta_t(n; X_{t-\Delta})$ depends on the yield curve factors as:

$$ \theta_t(n; X_{t-\Delta}) = a_t(n) + b_t(n) \text{Level}_{t-\Delta} + c_t(n) \text{Slope}_{t-\Delta} + d_t(n) \text{Curvature}_{t-\Delta} $$

Once the models are fit, the weights at time $t$ can be calculated as $\theta_t$ values
using the yield curve factors at $t$:

$$ 
w_t(1) = \theta_t(1; X_t) \\ 
w_t(5) = \theta_t(5; X_t) \\ 
w_t(10) = \theta_t(10; X_t)
$$

The entire procedure is wrapped into `model_multiplicative_regression()`, which,
consistent with the other `model_*()` functions developed so far, accepts a single
rolling split, along with the type of bullet (3 or 7-year), and returns the weights.

```{r, echo=FALSE}
model_multiplicative_regression <- function(split, bullet = "3") {
  
  # Training data. train_t is used for the final calculation of weights
  train <- analysis(split)
  train_t <- tail(train, 1)
  
  train_data <- train %>% 
    select(-maturity, -spot_rate) %>% 
    spread(maturity_nm, ER_n) %>%
    # The level/slope/curvature of last period 
    # is used to predict ER of this period
    mutate(level_1_month_ago     = lag(level),
           slope_1_month_ago     = lag(slope),
           curvature_1_month_ago = lag(curvature))
  
  # Construct the features of the regression
  lm_data <- train_data %>%
    transmute(
      ER_y              = !! as.name(bullet),
      ER_1              = `1`,
      ER_5              = `5`,
      ER_10             = `10`,
      ER_1_X_level      = `1`  * level_1_month_ago,
      ER_5_X_level      = `5`  * level_1_month_ago,
      ER_10_X_level     = `10` * level_1_month_ago,
      ER_1_X_slope      = `1`  * slope_1_month_ago,
      ER_5_X_slope      = `5`  * slope_1_month_ago,
      ER_10_X_slope     = `10` * slope_1_month_ago,
      ER_1_X_curvature  = `1`  * curvature_1_month_ago,
      ER_5_X_curvature  = `5`  * curvature_1_month_ago,
      ER_10_X_curvature = `10` * curvature_1_month_ago,
    )
  
  # Regression formula. -1 b/c no intercept
  form <- ER_y ~ ER_1  + ER_1_X_level  + ER_1_X_slope  + ER_1_X_curvature +
                 ER_5  + ER_5_X_level  + ER_5_X_slope  + ER_5_X_curvature +
                 ER_10 + ER_10_X_level + ER_10_X_slope + ER_10_X_curvature - 1
  
  # Run
  model <- lm(form, data = lm_data)
  
  # Coef tibble
  coefs <- enframe(coef(model)) %>% spread(name, value)
  
  # Use the weights and the level/slope/curve at time t to calc the weights
  weight_tbl <- coefs %>%
    bind_cols(train_t) %>%
    transmute(
      maturity_nm = list(c("1", "5", "10")),
      weight = list(c(
        ER_1  + ER_1_X_level  * level + ER_1_X_slope  * 
          slope + ER_1_X_curvature  * curvature,
        ER_5  + ER_5_X_level  * level + ER_5_X_slope  * 
          slope + ER_5_X_curvature  * curvature,
        ER_10 + ER_10_X_level * level + ER_10_X_slope * 
          slope + ER_10_X_curvature * curvature
      ))
    ) %>%
    unnest()
    
  weight_tbl
}
```

## Error Calculation

Hedging error for the weights set at time $t$ are calculated at time $t+\Delta$ as:

$$ \epsilon_{t+\Delta} = w_t(1) ER_{t+\Delta}(1) + w_t(5) ER_{t+\Delta}(5) + w_t(10) ER_{t+\Delta}(10) $$
These weights are then aggregated using RMSHE to determine overall model performance.

```{r, echo=FALSE}
hedging_error <- function(split, weight, bullet = "3") {
  test <- assessment(split)
  
  ER_model <- test %>% 
    filter(maturity_nm %in% c("1", "5", "10")) %>%
    left_join(weight, "maturity_nm") %>%
    summarise(ER_model = sum(ER_n * weight)) %>%
    pull(ER_model)
  
  ER_bullet <- test %>% 
    filter(maturity_nm %in% bullet) %>%
    pull(ER_n)
    
  tibble(error = ER_model - ER_bullet)
}
```

## Model Selection Interface

A practical way to call all of the above models would be through an interface
function that take as parameters: the data, the model to run, and the bullet to hedge against.
Such a function was developed, and returns a data frame of the model type, the
bullet used, the date of the error calculation, the weights, and the hedging errors.

```{r, echo=FALSE}
retrieve_modeling_function <- function(model) {
  switch(model,
   "modified_duration" = purrr::partial(model_duration, type = "modified"),
   "macaulay_duration" = purrr::partial(model_duration, type = "macaulay"),
   "regression"        = model_regression,
   "multiplicative_regression" = model_multiplicative_regression
  )
}
```

```{r, echo=FALSE}
apply_model <- function(.data, model, bullet) {
  modeling_fn <- retrieve_modeling_function(model)
  
  # Run the models
  .model_metrics <- .data %>%
    transmute(
      model  = model,
      bullet = bullet,
      
      # Extract the date of the error calculation
      error_date = map(
        .x = splits, 
        .f = ~assessment(.x) %>% 
          tail(1) %>% 
          select(date) %>% 
          rename(error_date = date)),
      
      # Run the model and return the weights
      weight = map(splits, ~modeling_fn(.x, bullet)),
      
      # Calculate the error at each modeling date
      error = map2(splits, weight, ~hedging_error(.x, .y, bullet))
    ) %>%
    unnest(error_date)

  .model_metrics
}
```

```{r, echo=FALSE}
root_mean_squared_hedging_error <- function(errors) {
  n <- length(errors)
  sqrt(1 / n * sum(errors ^ 2))
}
```

```{r, echo=FALSE}
summarise_total_hedging_error <- function(.data, error_col) {
  error_col <- enquo(error_col)
  
  # Calculate the RMSHE
  .data %>%
    unnest(!!error_col) %>%
    summarise(
      RMSHE  = root_mean_squared_hedging_error(!!error_col)
    )
}
```

## Model Application

```{r, echo=FALSE}
invokable <- tibble::tibble(
  f = "apply_model",
  params = list(
    list(.data = model_data, model = "modified_duration",         bullet = "3"),
    list(.data = model_data, model = "modified_duration",         bullet = "7"),
    list(.data = model_data, model = "macaulay_duration",         bullet = "3"),
    list(.data = model_data, model = "macaulay_duration",         bullet = "7"),
    list(.data = model_data, model = "regression",                bullet = "3"),
    list(.data = model_data, model = "regression",                bullet = "7"),
    list(.data = model_data, model = "multiplicative_regression", bullet = "3"),
    list(.data = model_data, model = "multiplicative_regression", bullet = "7")
  )
)

results <- invoke_map_dfr(invokable$f, invokable$params)
```

Finally, a data frame of function calls and parameter sets is created to easily iterate over all of the models. 
This data frame has a very compact form:

```{r}
invokable
```

The first column contains the function interface, `apply_model()`. Each element of the `params` column contains the data to be used in the model,
the model type, and the bullet to hedge against.

```{r}
invokable$params[[1]]
```

## Model Results

Table \@ref(tab:hedging-results) displays the RMSHE results from each model. The 
regression methods significantly outperformed the duration based models, with
much lower RMSHE. Among the duration models, Macaulay duration did marginally better,
but the statistical significance is likely negligible. Among the regression models,
the multiplicative regression did slightly better than the simple regression
with the 3-year bullet and slightly worse with the 7-year bullet. The results
are so close, however, that I am inclined to conclude that the simpler regression
model is the best and most parsimonious model of the four. There is little to indicate that longer maturity bullets are harder to hedge than shorter maturity bullets. In fact, with the regression models, the opposite seems
to be true.

```{r hedging-results, echo=FALSE, warning=FALSE}
results %>%
  group_by(model, bullet) %>%
  summarise_total_hedging_error(error) %>%
  ungroup() %>%
  mutate(model = recode(model, 
                        "macaulay_duration" = "Macaulay", 
                        "modified_duration" = "Modified", 
                        "multiplicative_regression" = "Multipl. Regression", 
                        "regression" = "Regression")) %>%
  rename(Model = model, Bullet = bullet) %>%
  knitr::kable(booktabs = TRUE, caption = "Hedging Model Results") %>%
  kableExtra::kable_styling(position = "center", latex_options = "HOLD_position")
```

```{r, echo=FALSE}
results_chartable <- results %>%
  mutate(model = recode(model, 
                        "modified_duration" = "Modified Duration", 
                        "macaulay_duration" = "Macaulay Duration", 
                        "regression" = "Regression",
                        "multiplicative_regression" = "Multiplicative Regression")) %>%
  unnest(weight) %>% 
  spread(maturity_nm, weight) %>%
  mutate(`Risk Free` = 1 - `1` - `10` - `5`) %>%
  gather("maturity_nm", "weight", -(model:error_date)) %>%
  mutate(maturity_nm = forcats::as_factor(maturity_nm) %>% forcats::fct_relevel("1", "5", "10", "Risk Free"))
```

Because a rolling model was fit, a similar analysis can be performed as Section \@ref(coef-stability) where
the stability of the coefficients over time was analyzed. In this case, the assignment of hedging weights by the models over time can be analyzed rather than just the coefficients. Some interesting insights can be gathered from Figure \@ref(fig:weight-stability). Keeping in mind that duration models are _only_ allowed to assign weight to a 1 year
and 10 year zero, it is interesting to see how stagnant the weighting is between
the 1 year and 10 year. The fact that the model essentially does not have the 
flexibility to vary the weights much over time likely contributes to its poor performance. The duration
model also never shorts, which likely helps the regression models. The weightings
in the duration models do make sense, with more than 50% being assigned to the
1-year bond to hedge the 3-year bullet, and more than 50% being assigned to the 10-year
the hedge the 7-year bullet. In the regression models, there is an implicit risk free instrument that weight
can be assigned to, so a 4th row in the figure is included to incorporate that. The
simple regression model only shifts weights around in periods of high stress, 2001
and the jump around 2008 being two examples, but even so, the weightings are surprisingly 
stable over time. This is especially apparent when contrasted with the multiplicative
regression. Weights are highly varied over time, but the performance gain for this
is limited as seen in the RMSHE results, questioning the need for the complex model.

```{r weight-stability, fig.width=10, echo=FALSE, fig.pos = 'H', fig.cap="Hedging weight assignment over time", warning=FALSE, fig.align='center', fig.height=8, cache=TRUE}
results_chartable %>%
  mutate(model = recode(model, 
                        "modified_duration" = "Modified Duration", 
                        "macaulay_duration" = "Macaulay Duration", 
                        "regression" = "Regression",
                        "multiplicative_regression" = "Multiplicative Regression")) %>%
  unnest(weight) %>% 
  spread(maturity_nm, weight) %>%
  mutate(`Risk Free` = 1 - `1` - `10` - `5`) %>%
  gather("maturity_nm", "weight", -(model:error_date)) %>%
  mutate(maturity_nm = forcats::as_factor(maturity_nm) %>% forcats::fct_relevel("1", "5", "10", "Risk Free")) %>%
  ggplot(aes(x = error_date, y = weight, color = bullet)) + 
  geom_line() + 
  facet_wrap(maturity_nm~model, ncol = 4) +
  theme_minimal() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "", y = "Weight", color = "Bullet")
```

Another way to view the weights over time is by stacking them. This is done in
Figure \@ref(fig:stacked-weight-stability) and provides another unique view into the 
change in the total weight distribution over time. This view confirms the stagnant
weights in the duration models, but also offers some other new insights. In the regression models, hedging
the 3-year bullet requires shorting the 10-year and the risk free, while hedging
the 7-year bullet only required hedging the 1-year bond. This view also further demonstrates
the wild swings in the multiplicative regression, for questionable performance gains.

```{r stacked-weight-stability, fig.width=10, echo=FALSE, fig.pos = 'H', fig.cap="Stacked weight assignment over time", warning=FALSE, fig.align='center', fig.height=8, cache=TRUE}
results_chartable %>%
  ggplot(aes(x = error_date, y = weight, fill = maturity_nm)) +
  geom_col() +
  facet_wrap(bullet ~ model, ncol = 4) +
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2") +
  labs(x = "", y = "Weight", fill = "Asset") +
  geom_hline(yintercept = 0, size = .1)
```





