--- 
title: "Finn 6211 - Final Project"
subtitle: "Davis Vaughan"
author: "mvaugh15@uncc.edu"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
classoption: openany
bibliography: [book.bib, packages.bib]
biblio-style: apalike
header-includes:
   - \usepackage{float}
link-citations: yes
description: "This is the final project of Finn 6211, with the intention of getting comfortable with spot rate data, their relation to yield curve factors, and various hedging strategies."
---

```{r, echo=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set("size" = "small")
```

# Introduction {#titlepage}

This is the final project of Finn 6211, with the intention of getting comfortable with spot rate data, their relation to yield curve factors, and various hedging strategies.

The entire analysis, including the code and the report, can be found at [https://github.com/DavisVaughan/final-project-book](https://github.com/DavisVaughan/final-project-book). An HTML version of the report is hosted at [http://final-project-finn-6211.davisvaughan.com/](http://final-project-finn-6211.davisvaughan.com/), and may be preferable to the PDF version, which can be found at [http://final-project-finn-6211.davisvaughan.com/final-project-book.pdf](http://final-project-finn-6211.davisvaughan.com/final-project-book.pdf).

An R package has been created to accompany the report. It contains a number of 
helper functions for cleaning data, manipulating the time series, and creating
the hedging strategies. The package is named `ratekit` and can be found on
GitHub at [https://github.com/DavisVaughan/ratekit](https://github.com/DavisVaughan/ratekit).

The structure of the report is as follows. Chapter 2 is dedicated to retrieving
and cleaning the data. Chapter 3 is focused on the calculation of features used
in the questions. Chapters 4, 5, and 6 answer the three questions required in
the report.

The code used to create the analysis is split into two places. In the `R/` folder
of the attached zip file are the files used for downloading the data, cleaning it, 
and creating the fixed income features. The actual analysis and answering of the 
questions is done inside the `.Rmd` files found in the top level of the zip file.

This report was written with [bookdown](https://bookdown.org/yihui/bookdown/), a
book authoring package for R.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
